---
title: "Home-range size across grid-cell sizes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Home-range size across grid-cell sizes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Grid-based home-range estimates depend on the spatial resolution of the grid. Here we examine how the estimated home-range size changes as grid-cell area increases.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gridHR)
```

```{r}
hr_cell_size(
  gibbons,
  min = 50,
  max = 10000,
  interval = 500,
  cell_shape = "hex",
  title = "Grid-based home-range of gibbons across hexagonal cell areas"
)
```

The figure shows that home-range size stabilizes at approximately 0.5 ha (5,000 square meters) hexagonal cell area, suggesting that larger cells provide little additional change in the estimated home-range size.
